\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-05-01 europa.tex
updates

\comment{I think we should expand on this section which, as the
  reviewer notes, the inter-relationships between the elements in Fig
  \ref{fig:europa-architecture} below is not fully captured. I think
  an illustrative example of how these components work together might
  be very useful. This might also help the sub-section end less
  abruptly than it is currently.}

@@

Describing problem domains and problem instances in terms of objects,
variables and domain rules in a programmatic way is cumbersome to the
extent of being impractical; this motivated the creation of modeling
languages that support those definitions in a much more natural and
compact way. \texttt{NDDL} (New Domain Description Language
[pronounced ``noodle'']) \cite{NDDL} is the main modeling language for
\eu and will be described in detail in the Modeling section
below. Another, \texttt{ANML} is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

Before diving into the details of each module, a simple example illustrates how they interact at a high level:

@@

The snippet above is a NDDL model of a light bulb and a switch that controls it (NDDL is covered in more detail below), along with a problem instance that requires the light bulb to be off by time=10, given an initial state with the bulb on at time=0. The NDDL module allows \eu to ingest this model and problem description. Once the NDDL statements are parsed, the different object types, object instances and constraints are instantiated inside the Plan Database and Constraint Engine modules where they are available for both the user and the solver module to query and manipulate programmatically. This instantiation of the initial state constitutes the initial plan that \eu holds for this particular problem instance. 
The user can then invoke functionality available in the solver module to modify or extend the initial plan in a way that eliminates any flaws (violated constraints, unsatisfied goals, oversubscribed resources) in the current plan. When actions are added to a plan, the Rules Engine module ensures that domain rules expressed in the model are enforced when \eu creates new action or predicate instances, for instance, if a LightSwitch::turnOn action is added to the plan, the RulesEngine  ensures that requirements for the existence of the conditions and effects of that action (that is, the corresponding LightBulb must be off before and on after the action) are posted through the PlanDatabase and ConstraintEngine modules.

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eue,
\texttt{NDDL} allows the user to specify the different components of a
problem domain in a precise and concise manner. Among the main
features that make \texttt{NDDL} a powerful tool for describing
problem domains are that it is object oriented. \texttt{NDDL} supports
classes and inheritance in similar fashion to popular Object-Oriented
languages like Java and C++. It also supports polymorphism for some of
the planning components. Using object classes and instances is a
time-tested approach to naturally describe problem domains
\cite{rumbaugh}. It offers declarative constructs to define
constraints and causality dependencies between actions and effects and
finally it offers procedural constructs to populate \eus \texttt{PDB}
with specific problem instances expressed in terms of objects,
variables, constraints, facts and goals. To articulate the most important
elements of \texttt{NDDL}, we show how the Shopping Agent problem
could be modeled:

@@

\item[\textbf{Object Types}:] \texttt{NDDL} supports an object-oriented
  approach, where common data and behavior can be abstracted as a
  class. In this case, we have the \textit{Product} class represent
  the products that the Shopping Agent needs, \textit{Location}
  represents the places where products can be found,
  \textit{ProductLocation} represents the association arising when a
  product can be purchased at a location. The Shopping Agent itself is
  represented by an \textit{Agent} class, which relies on the
  \textit{AgentLocation} class to keep track of where the agent is at
  any point in time. \texttt{NDDL} supports a single rooted class
  hierarchy so that state and behavior can be extended through
  inheritance. We take advantage of that feature by making
  \textit{AgentLocation} inherit from \eus built-in \textit{Timeline}
  class which doesn't allow predicates to overlap in time for the same
  object instance. This feature is used for instance, to ensure that
  an agent can only be at one place at a time in our example.

@@

Domain rules can be defined in \texttt{NDDL} in the context of a
predicate or an action. In the example above we can see the main
elements of a rule definition:

@@

An \eu user can introduce new constraint types by implementing
specialized constraint and propagator classes; however, \eu provides a
built-in constraint library with a range of arithmetic, temporal, set
and resource constraint types so that a large number of problem
domains can be modeled without having to introduce new constraint
types. Tables \ref{tab:calconst}, \ref{tab:setconst},
\ref{tab:objhierarchy} and \ref{tab:misc} show the main constraint
types included in \eus Constraint Library.

@@

Temporal and resource constraints are especially important to model
real-world planning domains.

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A Simple Temporal Network (STN)]{\label{fig:stn1}\includegraphics[width=0.4\columnwidth]{figs/stn-graph.pdf}}
  \subfloat[\small Resulting distance graph from \ref{fig:stn1}.]{\label{fig:stn2}\includegraphics[width=0.4\columnwidth]{figs/distance-graph.pdf}}
  \caption{\small An illustration of a STN and its corresponding
    distance graph.}
  \label{fig:stn}
\end{figure}

Most real-world planning problems need to deal with plans made up of
actions that typically depend upon limited resources (\eg fuel,
storage capacity, battery state of charge, data buffers).  \eu
provides built-in support for resource computation. In \eus
representation resource object types come in three forms with related
constraints. \textbf{Reservoir} resources can have both
\texttt{consume(resource,time,quantity)} and
\texttt{produce(resource,time,quantity)} constraints; production and
consumption is instantaneous. \textbf{Reusable} resources on the other
hand, have constraints that use a quantity of the resource for a
specific duration (\texttt{uses(resource,start,duration,quantity)}),
which are then subsequently released. \textbf {Unary} resources are
reusables with capacity$ = 1$, which allows \eu to perform run time
savings in comparison to the generic reusable class. Configuration
options are offered such that in addition to the usual constraint
violations when a resource is completely depleted, plan flaws can be
generated whenever a resource level is outside a specific upper or
lower bound, or when instantaneous production or consumption exceeds a
specific threshold.

Since \eu supports a flexible temporal representation, where time is
represented by intervals, instead of single values, determining
whether a resource is the source of a plan flaw or a constraint
violation in general is substantially more difficult than just
computing a profile of the resource and checking against some limits.
With temporal flexibility, a number of resource profiles could be
possible depending on precisely when each production/consumption event
occurs out of a range of possible values that are in the domain of the
variables that represent the temporal extent for the event.
\cite{Muscettola04,Muscettola06} proposed an algorithm where the
variables that represent the time for production/consumption events
are placed in a temporal network, similar to the network used by the
temporal propagator; constraints on the resource can then be seen as
quantities flowing through that network. Computing the maximum flow
then, yields the maximum resource usage and thus a lower bound on the
resource level; computing a minimum flow will similarly yield an upper
bound. Since maximum flow on a temporal network can be computed in
polynomial time \comment{definition needed?}, \eu takes advantage of
this algorithm to efficiently compute upper and lower bound envelopes
on the resource level. Fig. \ref{fig:resenvelopes} shows an example of
such a resource computation.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/europa-resource-envelopes.png}
\caption{\small An example of resource envelopes computed by \eu.}
\label{fig:resenvelopes}
\end{figure}

@@

\begin{algorithm}[H]
\KwIn{PartialPlan $plan$}
\KwOut{\texttt{true} if a plan solution was found, \texttt{false} otherwise}
\Begin{
  \lIf{isInconsistent(plan)}{\Return \texttt{false}\;}\label{li:propagate}
  \BlankLine
  Flaw $flaw \leftarrow chooseFlaw(plan)$ \; \label{li:selectflaw}
  \lIf{ $\emptyset = flaw$ }{\Return \texttt{true} \;} \label{li:noflaw}
  \BlankLine
  DecisionPoint $decision \leftarrow makeDecisionPoint(flaw, plan)$ \; \label{li:decisionpt}
  \While{ $decision.hasNext()$ }{ \label{li:decisionloop}
    PartialPlan $pp \leftarrow decision.executeNext()$ \; \label{li:decide}
    \lIf{ $solve(pp)$ }{ \Return \texttt{true} \; } \label{li:recurse}
    \lElse{ $decision.undo()$ \; } \label{li:backtrack}
   }
   \Return \texttt{false} \; \label{li:noplan}
 }
\caption{$\mathrm{bool} ~ solve(plan)$}
\label{alg:europa:solve}
\end{algorithm}

>>>>

@@

Describing problem domains and problem instances in terms of objects,
variables and domain rules in a programmatic way is cumbersome to the
extent of being impractical; this motivated the creation of modeling
languages that support those definitions in a much more natural and
compact way. \nd (New Domain Description Language
[pronounced ``noodle'']) \cite{NDDL} is the main modeling language for
\eu and will be described in detail in the Modeling section
below. Another, \texttt{ANML} is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

\kcomment{Before going into the details of each module, a simple
  example illustrates how they interact at a high level:}

@@

\kcomment{The snippet above is a \nd (covered in more detail in
  Section \ref{sec:europa:modeling}) model of a light bulb and a
  switch that controls it, along with a problem instance that requires
  the light bulb to be off by time$=10$, given an initial state with
  the bulb on at time$=0$. The \nd module allows \eu to ingest this
  model and the problem description. Once the \nd statements are
  parsed, different object types, object instances and constraints are
  instantiated inside the \texttt{PDB} and \texttt{CRE} modules where
  they are available for both the user and the solver module to query
  and manipulate programmatically. This instantiation of the initial
  state constitutes the initial plan that \eu holds for this
  particular problem instance.

  The user can then invoke functionality available in the solver
  module to modify or extend this initial plan in a way that
  eliminates any flaws (violated constraints, unsatisfied goals,
  oversubscribed resources). When actions are added to a plan, the
  \texttt{RE} module ensures that domain rules expressed in the model
  are enforced when \eu creates new action or predicate instances. For
  instance, if a \texttt{LightSwitch::turnOn} action is added to the
  plan, the \texttt{RE} ensures that requirements for the existence of
  the conditions and effects of that action (that is, the
  corresponding \texttt{LightBulb} must be off before and on after the
  action) are posted through the \texttt{PDB} and \texttt{CRE}
  modules.  }

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eue,
\nd allows the user to specify the different components of a problem
domain in a precise and concise manner. Among the main features that
make \nd a powerful tool for describing problem domains are that it is
object oriented. \nd supports classes and inheritance in similar
fashion to popular Object-Oriented languages like Java and C++. It
also supports polymorphism for some of the planning components. Using
object classes and instances is a time-tested approach to naturally
describe problem domains \cite{rumbaugh}. It offers declarative
constructs to define constraints and causality dependencies between
actions and effects and finally it offers procedural constructs to
populate \eus \texttt{PDB} with specific problem instances expressed
in terms of objects, variables, constraints, facts and goals. To
articulate the most important elements of \nde, we show how
the Shopping Agent problem could be modeled:

@@

\item[\textbf{Object Types}:] \nd supports an object-oriented
  approach, where common data and behavior can be abstracted as a
  class. In this case, we have the \textit{Product} class represent
  the products that the Shopping Agent needs, \textit{Location}
  represents the places where products can be found,
  \textit{ProductLocation} represents the association arising when a
  product can be purchased at a location. The Shopping Agent itself is
  represented by an \textit{Agent} class, which relies on the
  \textit{AgentLocation} class to keep track of where the agent is at
  any point in time. \nd supports a single rooted class
  hierarchy so that state and behavior can be extended through
  inheritance. We take advantage of that feature by making
  \textit{AgentLocation} inherit from \eus built-in \textit{Timeline}
  class which doesn't allow predicates to overlap in time for the same
  object instance. This feature is used for instance, to ensure that
  an agent can only be at one place at a time in our example.

@@

Domain rules can be defined in \nd in the context of a predicate or an
action. In the example above we can see the main elements of a rule
definition:

@@

An \eu user can introduce new constraint types by implementing
specialized constraint and propagator classes; however, \eu provides a
built-in constraint library with a range of arithmetic, temporal, set
and resource constraint types so that a large number of problem
domains can be modeled without having to introduce new constraint
types. Tables \ref{tab:calconst}, \ref{tab:setconst},
\ref{tab:objhierarchy} and \ref{tab:misc} show the main constraint
types included in \eus Constraint Library.  Temporal and resource
constraints are especially important to model real-world planning
domains.

@@

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A Simple Temporal Network (STN). \kcomment{Nodes correspond
  to variables and their domains and arc's the \emph{constraint} or
  distance between them.}]{\label{fig:stn1}\includegraphics[width=0.4\columnwidth]{figs/stn-graph.pdf}}
  \subfloat[\small The resulting distance graph from \ref{fig:stn1}.]{\label{fig:stn2}\includegraphics[width=0.4\columnwidth]{figs/distance-graph.pdf}}
  \caption{\small An illustration of a STN and its corresponding
    distance graph.}
  \label{fig:stn}
\end{figure}

Most real-world planning problems need to deal with plans made up of
actions that typically depend upon limited resources (\eg fuel,
storage capacity, battery state of charge, data buffers etc).  % \eu
% provides built-in support for resource computation.
In \eus representation, resource object types come in three forms with
related constraints. \textbf{Reservoir} resources can have both
\texttt{consume(resource,time,quantity)} and
\texttt{produce(resource,time,quantity)} constraints; production and
consumption is instantaneous. \textbf{Reusable} resources on the other
hand, have constraints that use a quantity of the resource for a
specific duration (\texttt{uses(resource,start,duration,quantity)}),
which are then subsequently released. \textbf {Unary} resources are
reusables with capacity$ = 1$, which allows \eu to perform run time
savings in comparison to the generic reusable class. Configuration
options are offered such that in addition to the usual constraint
violations when a resource is completely depleted, plan flaws can be
generated whenever a resource level is outside a specific upper or
lower bound, or when instantaneous production or consumption exceeds a
specific threshold.

Since \eu supports a flexible temporal representation, where time is
represented by intervals, instead of singletons, determining whether a
resource is the source of a plan flaw or a constraint violation in
general is substantially more difficult than just computing a profile
of the resource and checking against some limits.  With temporal
flexibility, a number of resource profiles could be applicable
depending on precisely when each production/consumption event occurs
out of a range of possible values that are in the domain of the
variables that represent the temporal extent for the event.
\cite{Muscettola04,Muscettola06} proposed an algorithm where the
variables that represent the time for production/consumption events
are placed in a temporal network, similar to the network used by the
temporal propagator; constraints on the resource can then be seen as
quantities flowing through that network. Computing the maximum flow
then, yields the maximum resource usage and thus a lower bound on the
resource level; computing a minimum flow will similarly yield an upper
bound. Since maximum flow on a temporal network can be computed in
polynomial time \cite{corman}, \eu takes advantage of this algorithm
to efficiently compute upper and lower bound envelopes on the resource
level. Fig. \ref{fig:resenvelopes} shows an example of such a resource
computation.

\begin{figure}
\centering
\includegraphics[scale=0.2]{figs/europa-resource-envelopes.png}
\caption{\small An example of resource envelopes computed by \eu.}
\label{fig:resenvelopes}
\end{figure}

@@

\begin{algorithm}[H]
\KwIn{PartialPlan $plan$}
\KwOut{\texttt{true} if a plan solution was found, \texttt{false} otherwise}
\Begin{
  \lIf{isInconsistent(plan)}{\Return \texttt{false}\;}\label{li:propagate}
  \BlankLine
  Flaw $flaw \leftarrow chooseFlaw(plan)$ \; \label{li:selectflaw}
  \lIf{ $\emptyset = flaw$ }{\Return \texttt{true} \;} \label{li:noflaw}
  \BlankLine
  DecisionPoint $decision \leftarrow makeDecisionPoint(flaw, plan)$ \comment{\{\emph{potential branching point(s)}\}}\;\label{li:decisionpt}
  \While{ $decision.hasNext()$ }{ \label{li:decisionloop}
    PartialPlan $pp \leftarrow decision.executeNext()$ \; \label{li:decide}
    \lIf{ $solve(pp)$ }{ \Return \texttt{true} \; } \label{li:recurse}
    \lElse{ $decision.undo()$ \comment{\{\emph{possible backtracking point}\}}\; } \label{li:backtrack}
   }
   \Return \texttt{false} \; \label{li:noplan}
 }
\caption{$\mathrm{bool} ~ solve(plan)$\kcomment{: To deliberate and
    generate partial plans}}
\label{alg:europa:solve}
\end{algorithm}
